---
title: "Assignment 1 - Group 1"
output: pdf_document
---
# Introduction
This dataset contains a list of video games with sales greater than 100,000 copies from Kaggle. The sales numbers in the dataset are in millions. There are 16,598 records in the dataset and the data was last updated in 2016. This dataset will be used to perform data analysis for the purpose of Assignment 1.


# Load the necessary packages
```{r echo=TRUE}
library(ggplot2)
library(dplyr)
library(tidyverse)
library (readr)
```

# Load dataset
Because the file location will be different for everyone, we load the dataset directly from Github raw file

```{r echo=TRUE}
urlfile="https://raw.githubusercontent.com/trngminhtrang/DataAnalysis--Video-Games-Sales--Historical-Data/main/vgsales.csv"

vgsales <-read_csv(url(urlfile))
```

# Print the structure of the dataset
```{r echo=TRUE}
str(vgsales)
```

# List the variables in the dataset
```{r echo=TRUE}
ls(vgsales)
```

# Print the top 15 rows of the dataset
```{r echo=TRUE}
head(vgsales,15)
```
# Write a user defined function
User defined function “model” calculates the sum of two variables namely  NA_Sales and EU_Sales from the data set and stores it into new variable called "sum" in vgsales dataset

```{r echo=TRUE}
model<-function(x,y){x+y}
vgsales$sum = model(vgsales$NA_Sales, vgsales$EU_Sales)
head(vgsales$sum)
```

# Using filter command to filter out sales where Global_Sales are > 10
```{r echo=TRUE}
vgsalesnew2 = as.data.frame(vgsales %>% filter(Global_Sales > 10))
summary(vgsalesnew2)
```

# Identify the dependent & independent variables and create a new data frame by joining these variables
As Global_Sales  the total sales worldwide, which is the number of all regions sales combined, we can identify it as dependent variable. We select NA_Sales as the independent variable for this task. In this case, we extract the 1st & 6th coloumns and create a new data frame called "vgsalesnew1"

```{r echo=TRUE}
vgsalesnew1 = as.data.frame(vgsales %>% select(7,11))
summary(vgsalesnew1)
```

# Remove missing values
## Replace N/A values in the dataset with NA
We notice that the missing values in the dataset were recorded as N/A, which appears to not be treated as NA. We will replace the N/A values with NA so R can recognize the missing values

```{r echo=TRUE}
vgsales[vgsales=="N/A"]=NA
```

## Identify the number of missing values
```{r echo=TRUE}
sum(is.na(vgsales))
```

## Remove missing values from the dataset
```{r echo=TRUE}
vgsales1 <- vgsales[complete.cases(vgsales), ]
str(vgsales1)
```

There are 16,291 records in the new data frame while the original dataframe has 16,598 records. This means 329 missing values have been removed from the new dataframe.

# Identify and remove duplicated data

## Identify duplicated data throughout the dataset

```{r echo=TRUE}
duplicate <- duplicated(vgsales)
sum(duplicate)
```

There is no duplicated rows in the dataset. We will remove duplicated data in the EU_Sales variable. 

## Identify and remove duplicated data in EU_Sales variable
Identify the number of duplicated  in EU_Sales
```{r}
sum(duplicated(vgsales$EU_Sales))
```

Remove duplicated data

```{r echo=TRUE}
noduplicate_EU <- vgsales %>% distinct(Distinct_EU = EU_Sales)
noduplicate_EU
```
There are 305 distinct records from the new dataset noduplicate_EU. This means 16,293 duplicated data has been removed from EU_Sales.

# Reorder multiple rows in descending order
We reorder the dataset in descending order of the EU_Sales variable

```{r echo=TRUE}
vgsales %>% arrange(desc(EU_Sales))
```
# Rename some column names in the dataset
Rename the "Rank" column with "Ranking" and "Name" column with "Games".

```{r echo=TRUE}
names(vgsales)[1] <- 'Ranking'
names(vgsales)[2] <- 'Games'
head(vgsales)
```
# Add new variables by using a mathematical function
```{r echo=TRUE}
vgsales$New_JP_Sales = vgsales$JP_Sales*2
head(vgsales)
```
# Create a training set using random number generator engine

```{r echo=TRUE}
set.seed(1)
vgsales%>% sample_n (15, replace = FALSE)
```

# Print the summary statistics of the dataset
```{r echo=TRUE}
summary(vgsales)
```

# Perform statistical functions using EU_Sales varable

## Calculate Mean
```{r echo=TRUE}
mean(vgsales$EU_Sales)
```

## Calculate Median
```{r echo=TRUE}
median(vgsales$EU_Sales)
```

## Calculate Mode
As R does not have a standard built-in function to calculate mode, we create a user function to calculate mode of EU_Sales in the dataset.

### Create the function to calculate Mode
```{r echo=TRUE}
getmode <- function(v) {
     uniqv <- unique(v)
     uniqv[which.max(tabulate(match(v, uniqv)))]
}
```

### Assign v to the EU_Sales variable of the dataset
```{r echo=TRUE}
v <- vgsales$EU_Sales
```

### Create a variable to store the Mode result
```{r echo=TRUE}
mode <- getmode(v)
```

### Print the Mode result
```{r echo=TRUE}
print(mode)
```
The mode value of EU_Sales is 0 which could mean that a majority of games were not available or were not released in EU.

## Calculate Range
```{r}
range(vgsales$EU_Sales)
```


# Plot a scatter plot for Year and EU_Sales variables
```{r echo=TRUE}
ggplot(data = vgsales, aes(Year, EU_Sales)) + geom_point(color = "blue") + theme(axis.text.x = element_text (angle=45, hjust=1, vjust=1)) + xlab("Year") + ylab("EU Sales") + ggtitle("Year vs EU Sales") + theme(plot.title = element_text(hjust = 0.5))
```

# Plot a bar plot for any two variables in  dataset

## Filter data for the last 10 years

We filter the dataset by the last 10 years to reduce the data size and make it more relevant for analysis. Since the dataset was last updated in 2016, we filter the time range from 2006 to 2016.

```{r echo=TRUE}
filtered_years <- filter(vgsales, Year >= 2006, Year<=2016)
```

The filtered dataset is stored in "filtered_years".

## Plot a bar plot for Year and Global Sales variables

This bar plot shows the global sales of videos games by genre over the last 10 years from the last year the dataset was updated (2016).

```{r echo=TRUE}
ggplot(data = filtered_years, aes(Year, Global_Sales)) + geom_bar(aes(fill=Genre), stat = "identity") + ylab("Global Sales") + ggtitle("Video games sales worldwide by genre from 2006-2016") + theme(plot.title = element_text(hjust = 0.5))
```

# Find the correlation between NA_Sales and Global_Sales
## Plot a scatter plot for NA_Sales and Global_Sales with a regression line
This scatter plot aims to find a correlation between the video games sales in the North America and the sales worldwide.

```{r echo=TRUE}
ggplot(data = filtered_years, aes(NA_Sales, Global_Sales))+geom_point(color="steelblue")+geom_smooth(aes(color = "Regression Line"), method="lm", se = FALSE) + scale_colour_manual(name="Legend", values = c("red")) + xlab("North America Sales") + ylab("Global Sales") + ggtitle("Correlation between NA Sales and Global Sales") + theme(plot.title = element_text(hjust = 0.5))
```

## Define X and Y variables for regression model

For the regression model, we are using the full dataset instead of the filtered dataset above to have an overall view

```{r echo=TRUE}
Y = vgsales$NA_Sales
X= vgsales$Global_Sales
```

## Find the correlation between NA_Sales (Y) and Global_Sales (X)

```{r echo=TRUE}
corrl = cor(X, Y,method = "pearson")
corrl
```

# Conclusion of Analysis

The global sales of a video game had reached as high as 82.7400 and as low as 0.01. The minimum sales of all regions are 0, and NA has the highest maximum sales of 41.49 while JP has the lowest maximum sales of 10.22.

From the "Video Games Sales Worldwide by Genre from 2006-2016" bar plot, we can see that Sports genre had generated the most sales in 2006 and remained the one of top genres in 2 years before it started to lose traction since 2010. Over the time, Action genre has remained the most popular genre since 2007 while Strategy games made the lowest to zero sales. Further research is needed to determine whether the game publishers dropped Strategy games out of their roadmap or they still rolled out new Strategy games but were unsuccessfully to make revenue from it.

For the "Correlation between NA Sales and Global Sales" scatter plot, it seems to be a correlation between these two variables. Moreover, the correlation coefficient value of 0.9410474 is very close to 1 which means that the NA_Sales and Global_Sales variables have a positive correlation. An increase in NA_Sales will be likely to generate an increase in the Global_Sales to a respective extent, which also means NA_Sales variable is made up a very high percentage of the Global_Sales value.

